
---
title: "Capstone data analysis"
author: "Yunxia Wei"
date: "2023-10-25"
output: 
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(readxl)
library(tidyverse)
```


#Load all States Dataset
```{r, warning=FALSE}
data_frames <- list()
file_paths <- list.files(path = "/Users/ywei/Desktop/Fall 2023/Capstone/Post-Secondary Employment Outcomes/Working folder/Wage Data/", pattern = "\\.xlsx$", full.names = TRUE)

for (file_path in file_paths) {
  df <- read_excel(file_path, skip = 5)  # Skip first 5 rows
  data_frames <- append(data_frames, list(df))
}
```

#Merge data files together
```{r}
combined_data <- do.call(rbind, data_frames) %>%
            rename("institution_code" = "institution",
                   "institution_name" = "label_institution",
                   "institution_degreelevel" ="label_degree_level")%>%
            mutate(institution_code = sub("^0+", "", institution_code))
```

#Load data from ipeds and rename variables
```{r}
ipeds_data <- read.csv("IPEDS Data/CSV_9252023-880.csv", stringsAsFactors = FALSE)%>%
            rename("institution code" = "HD2022.Office.of.Postsecondary.Education..OPE..ID.Number",
                   "longitude" = "HD2022.Longitude.location.of.institution",
                   "latitude" = "HD2022.Latitude.location.of.institution",
                   "12 month full-time enrollment" = "DRVEF122022.12.month.full.time.equivalent.enrollment",
                   "completers total" = "C2022_B.Grand.total",
                   "completers men" = "C2022_B.Grand.total.men",
                   "completers women" = "C2022_B.Grand.total.women",
                   "tuition and fees" = "DRVIC2022.Tuition.and.fees..2022.23",
                   "institution size category" = "HD2022.Institution.size.category",
                   "State abbreviation" ="HD2022.State.abbreviation",
                   "Bureau of Economic Analysis regions" ="HD2022.Bureau.of.Economic.Analysis..BEA..regions",
                   "Level of institution" = "HD2022.Level.of.institution",
                   "Control of institution" = "HD2022.Control.of.institution",
                   "Degree of urbanization" = "HD2022.Degree.of.urbanization..Urban.centric.locale.",
                   "Degree granting status" = "HD2022.Degree.granting.status")
```

#Limit the selection(from wage_data) to institutions that are present in the IPEDS dataset
```{r}
wage_data_for_use <-combined_data%>%
            mutate(in_ipeds = ifelse(institution_code%in%ipeds_data$`institution code`, "Yes","No"))%>%
            filter(in_ipeds == "Yes")
```

#Limit the selection(from ipeds_data) to institutions that are present in the wage dataset
```{r}
ipeds_data_for_use <-ipeds_data%>%
            mutate(in_wagedata = ifelse(`institution code`%in%wage_data_for_use$institution_code,"Yes","No"))%>%
            filter(in_wagedata == "Yes")%>%
            filter(check.duplicates == "keep")#The flag added in the CSV file identifies duplicate IDs with blank information
```

#Compare the wage data for use file and the IPEDS data for use file to verify if they have the same count of unique institutions
```{r}
unique_count_wage <- wage_data_for_use %>%
            summarise(unique_institution_count = n_distinct(institution_code))
unique_count_ipeds <- ipeds_data_for_use %>%
            summarise(unique_institution_count = n_distinct(`institution code`))

unique_count_wage
unique_count_ipeds
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
